---
title: "Lab 01 Report"
author: "YOUR NAME HERE"
date: "DATE HERE"
output:
  html_document:
    df_print: paged
---

# Skills

```{r load_data, warning = F, message = F}
# load package of helpful datasets and functions
library(tidyverse)

# load sleep dataset 
data("msleep")

?msleep

# print out the first few rows
head(msleep)
```

Below is an example code chunk. Use chunks like these to run the Lab 01 code. You do not need to turn anything in for the Skills section of Lab 01.

```{r}
# YOUR CODE HERE
```

# Practice and extension.

The ````ChickWeight```` dataset provides data on a sample of fifty chicks, each fed one of four different diets. For each chick, the data frame includes weight measurements at birth and every other day thereafter. First, load the data.

```{r load_chick_data, warning = F, message = F, include = F}
# YOUR CODE HERE
```

1. Print out the weights of the chicks on day 0.

```{r}
# YOUR CODE HERE
```

2. The last day of the experiment is Day 21. Print out the ending weights of the chicks.

```{r}
# YOUR CODE HERE
```

3. Calculate the mean ending weight of the chicks.
```{r}
# YOUR CODE HERE
```

4. Next, print out the starting weights of the chicks that ate Diet 1.

```{r}
# YOUR CODE HERE
```

5. Print out the ending weights of the chicks that ate Diet 1.

```{r}
# YOUR CODE HERE
```

6. For each diet, calculate the mean ending weight of the chicks that ate that diet. Is this a good way to compare the diets?

```{r}
# YOUR CODE HERE
```

7. Make a histogram of all of the chicks' birthweights. Use the x-axis label "Birth Weights" and the title "Frequency of Birthweights". **Provide a one-sentence summary of the plot.**

```{r}
# YOUR CODE HERE
```

8. Make a scatterplot with weight on the y-axis, and time on the x-axis. Use an appropriate title and axis labels, and change the color and shape of the points from their default values. **Provide a one-sentence summary of the plot. Additionally, give one reason this is a good way to plot this data, and one reason this plot may be incomplete.**

```{r}
# YOUR CODE HERE
```

9. Finally, use the pipe operator ```%>%``` and the ```filter``` function to create a new data frame ```diet_one``` that includes only the chicks eating Diet 1. 

```{r}
# YOUR CODE HERE
```

10. Use ```ggplot2``` to generate a line plot of the weights of the chicks that ate Diet 1 over time. **Provide a one-sentence summary of the resulting plot.**

```{r}
# YOUR CODE HERE
```

11. Generate your own visualization of the ```ChickWeight``` data. You can use any of the functions we've used above, but you are encouraged to think about other kinds of plots and figure out how to make them in R. **Provide a one-sentence summary of your plot. Additionally, give one reason this is a good way to plot this data, and one reason your plot may be incomplete.**
```{r}
# YOUR CODE HERE
```

12. Which diet produces the most weight gain? If you're not sure how to compute this in R, that is okay. If you can describe the set of steps you would take to compare the diets, that will be enough.