---
title: "Lab 02"
subtitle: "*Descriptive Statistics II*"
author: ""
institute: ""
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["style.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
 
---     
class: middle

In this lab, we will continue practice with descriptive statistics and making plots in R for a dataset on college tuition and pay.

---
class: middle

Today's data comes from the [`tidytuesday` project](https://github.com/rfordatascience/tidytuesday), which makes datasets freely available for practice with data visualization and exploration.

The data we'll be looking at today come originally from the Department of Education, but are summarized on [Tuitiontracker.org](https://www.tuitiontracker.org/). For more info on the cleaned dataset, see [this link](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-03-10/readme.md).

Other teachers and statisticians often use `tidytuesday` data to teach
 data analysis; you can find some examples [here](https://juliasilge.com/blog/tuition-resampling/) and [here](https://jonnylaw.rocks/blog/tidy-tuesday-us-tuition-data/).
---
class: middle

First, we need to load our dataset. We also load the `tidyverse` package to help us with data analysis.

```{r load_data, warning = F, message = F}
# load package of helpful datasets and functions
library(tidyverse)

# load data
tuition_cost <-
  readr::read_csv(paste0(
    'https://raw.githubusercontent.com/',
    'rfordatascience/tidytuesday/master/',
    'data/2020/2020-03-10/tuition_cost.csv'))
```
---
class: middle

#### 1. Use the `head` command to display the first few rows of data.

Each row corresponds to a school and degree type and provides the cost of in-state and out-of-state tuition.

```{r}
head(tuition_cost)
```
---
class: middle

#### 2. Use the "pipe" `%>%` operator with the `filter` command to select only schools in Washington.

```{r}
# save results in new object tuition_WA
tuition_WA <- tuition_cost %>%
  filter(state_code == "WA")

# print first few rows
head(tuition_WA) 
```
---
class: middle

#### 3. Use the `hist` command to produce a histogram of `in_state_total`, the **annual** in-state total costs for schools in Washington. Use appropriate `xlab` and `main` labels.

```{r, fig.height= 5, fig.width=7, fig.align = "center"}
hist(tuition_WA$in_state_total,
     xlab = "Total Costs",
     main = "WA College Annual In-State Total Costs")
```
---

class: middle

#### 4. Let's compare in-state total costs for public and private WA schools. What do you observe?

```{r, fig.height= 4, fig.width=8, fig.align = "center"}
public_WA <- tuition_WA %>%
  filter(type == "Public")
private_WA <- tuition_WA %>%
  filter(type == "Private")
par(mfrow = c(1, 2))
hist(public_WA$in_state_total,
     xlab = "Total Costs",
     main = "Public")
hist(private_WA$in_state_total,
     xlab = "Total Costs",
     main = "Private")
```
---

#### 5. Instead of using histograms, we can use boxplots to compare costs for public and private WA schools. What do you observe?

```{r, fig.height= 5, fig.width=8, fig.align = "center"}
boxplot(tuition_WA$in_state_total ~ # response variable
          tuition_WA$type, # explanatory variable
        horizontal = T, 
        xlab = "In-state Total Costs",
        main = "WA College Annual In-state Total Costs")
```
---
class: middle

#### 6. We can also split up the data by `degree_length`. Why are some of the boxplots missing?

```{r, fig.height= 4.5, fig.width=8.5, fig.align = "center"}
# change margins of plot to make x-axis visible
par(mar = c(5, 8, 3, 1)) 
boxplot(tuition_WA$in_state_total ~ # response variable
          tuition_WA$type + # explanatory variable 1
          tuition_WA$degree_length, # explanatory variable 2
        horizontal = T, las=1, # make boxes and labels horizontal
        xlab = "In-state Total Costs",
        main = "WA College Annual In-state Total Costs")
```
---
class: middle

#### 7. Let's repeat this plot but for all data:

```{r, fig.height= 4.5, fig.width=8.5, fig.align = "center"}
# change margins of plot to make x-axis visible
par(mar = c(5, 8, 3, 1)) 
boxplot(tuition_cost$in_state_total ~ # response variable
          tuition_cost$type + # explanatory variable 1
          tuition_cost$degree_length, # explanatory variable 2
        horizontal = T, las=1, # make boxes and labels horizontal
        xlab = "In-state Total Costs",
        main = "US College Annual In-state Total Costs")
```
---

class: middle



#### 8. The `other` categories don't seem to have much data. Let's remove those rows.

```{r, fig.height= 4.5, fig.width=8.5, fig.align = "center"}
plot_data <- tuition_cost %>%
  filter(type != "Other") %>%
  filter(degree_length != "Other")
```
---

class: middle

#### 9. Now, let's remake our boxplots. What conclusions can you draw?

```{r, fig.height= 4.5, fig.width=8.5, fig.align = "center"}
par(mar = c(5, 8, 3, 1)) 
boxplot(plot_data$in_state_total ~ # response variable
          plot_data$type + # explanatory variable 1
          plot_data$degree_length, # explanatory variable 2
        horizontal = T, las=1, # make boxes and labels horizontal
        xlab = "In-state Total Costs",
        main = "US College Annual In-state Total Costs")
```
---
class: middle

#### 10. Let's also make a two-way table to see how many schools of each type we have in our dataset.

```{r, fig.height= 4.5, fig.width=8.5, fig.align = "center"}
table(plot_data$type, plot_data$degree_length)
```
---

class: middle

#### 11. Finally, let's calculate some summary statistics:

```{r}
mean(plot_data$in_state_total)
mean(plot_data$out_of_state_total)
sd(plot_data$in_state_total)
sd(plot_data$out_of_state_total)
```
---
class: middle

These statistics are not very meaningful since we know that we have different groups (public vs. private vs. for-profit) within our dataset.

#### 12. Use the `group_by` and `summarize` commands to calculate **group-specific** means for each type of college (public, private, for-profit).

```{r}
plot_data %>%
  group_by(type) %>%
  summarize(mean(in_state_total)) %>%
  ungroup()
```
---
class: middle

We can also group by more than one variable.

#### 13. Calculate **group-specific** means for each type of college (public, private, for-profit) and each degree length (2-year, 4-year).

```{r}
plot_data %>%
  group_by(type, degree_length) %>%
  summarize(mean(in_state_total)) %>%
  ungroup()
```
---
class: middle

#### 14. Calculate the maximum, minimum, median, first quartile, and third quartile of the `in_state_total` data.

```{r}
max(plot_data$in_state_total)
min(plot_data$in_state_total)
median(plot_data$in_state_total)
quantile(plot_data$in_state_total, probs = .25)
quantile(plot_data$in_state_total, probs = .75)
```
---
class: middle

#### 15. Here is code for calculating **group-specific** medians for each type of college (public, private, for-profit) and each degree length (2-year, 4-year).

```{r}
plot_data %>%
  group_by(type, degree_length) %>%
  summarize(median(in_state_total)) %>%
  ungroup()
```
---

#### 16. Finally, here is code for calculating **group-specific** medians for each type of college (public, private, for-profit) and each degree length (2-year, 4-year).

---
class: middle

Making plots in `ggplot2` is fundamentally different from making plots in base R. Let's review the base R command. We only call one function, `hist`, which we provide with data (`tuition_WA$in_state_total`) and `xlab` and `main` arguments.

```{r, eval = F}
hist(tuition_WA$in_state_total,
     xlab = "Total Costs",
     main = "WA College In-State Total Costs")
```
---
class: middle
`ggplot2` plots are built gradually. We first specify the data for the plot (`ggplot(tuition_WA)`) and then tell the computer what kind of plot we want, (`geom_histogram()`). Within the `geom_histogram` command, we specify which data variables we want to use for which axes (`aes(x = in_state_total`)). We finally add on axis labels and titles.

```{r, eval = F}
ggplot(tuition_WA) +
  geom_histogram(aes(x = in_state_total), bins = 7) +
  xlab("Total Costs") +
  ggtitle("WA College In-State Total Costs")
```
