---
title: "Lab 03"
subtitle: "*Distributions*"
author: ""
institute: ""
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["style.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
 
---     
class: middle

In this lab, we'll practice working with different distributions in `R`.\pause

First, let's take a look at binomial distributions.
---
class: middle

#### Binomial distribution

A **binomial random variable** describes the number of successes in a fixed number of independent trials.


The **binomial distribution** describes the probability of having exactly $k$ successes in $n$ independent Bernoulli trials with probability of success $p$.

The probability of observing exactly $k$ successes in $n$ independent trials is given by
$$\binom{n}{k}p^{k}(1-p)^{n-k}=\frac{n!}{k!(n-k)!}p^{k}(1-p)^{n-k}$$
---
class: middle

#### Sampling from a binomial distribution

Recall that every binomial distribution has a *size* parameter which is the number of trials and a *probability of success* parameter. Below, we sample ten coin flips using the `rbinom` function

```{r}
coin_flips <- rbinom(n = 10, # number of samples
                     size = 1, # size parameter
                     prob = 0.5) # probability of success
coin_flips
```


---
class: middle

#### Sampling from a binomial distribution

We obtain more samples by changing $n$:

```{r}
coin_flips <- rbinom(n = 20, # number of samples
                     size = 1, # size parameter
                     prob = 0.5) # probability of success
coin_flips
```

---
class: middle

#### Sampling from a binomial distribution

Now, consider an experiment where we flip a coin twenty times. The below code runs this experiment one thousand times for a fair coin and then for an unfair coin.

```{r, cache = T}
fair_flips <- rbinom(n = 1000, # number of experiments
                     size = 20, # size parameter
                     prob = 0.5) # probability of success
unfair_flips <- rbinom(n = 1000, # number of experiments
                       size = 20, # size parameter
                       prob = 0.75) # probability of success
```

---
class: middle

#### Sampling from a binomial distribution

Now, let's compare our results:

```{r, fig.height = 5, fig.width = 8, fig.align = "center"}
par(mfrow = c(1, 2)) # combine plots into one row with two columns
hist(fair_flips, main = "Fair", xlim = c(0, 20), breaks = 10)
hist(unfair_flips, main = "Unfair", xlim = c(0, 20), breaks = 10)
```

---
class: middle
#### Probability mass function

For a **discrete** random variable $X$, we usually can specify **distribution** by just giving the probabilities of each possible value of $X$.

We call the function $P(X=k)$ the **probability mass function**.

**ex.**

$$P(X=k)=\binom{n}{k}p^{k}(1-p)^{n-k}=\frac{n!}{k!(n-k)!}p^{k}(1-p)^{n-k}$$
---
class: middle
#### Calculating probabilities for a binomial distribution

The `dbinom` command gives the **probability mass function**; `dbinom` calculates the probability of observing `x` successes out of `size` independent trials, each with probability of success `prob`:

```{r}
dbinom(x = 0, # number of successes
       size = 1, # size parameter
       prob = 0.5) # probability of success
```
```{r, echo = F, fig.height = 5, fig.width = 8, fig.align = "center"}
n <- 1
P <- 0.5
data <- dbinom(x=0:n,size=n, prob=P)
names(data) <- 0:n
cols <- rep("grey", n + 1)
cols[1] <- "tomato"
barplot(data, col = cols, main = "x = 0, n = 1, p = 0.5")
```
---
class: middle
#### Calculating probabilities for a binomial distribution

The `dbinom` command gives the **probability mass function**; `dbinom` calculates the probability of observing `x` successes out of `size` independent trials, each with probability of success `prob`:

```{r}
dbinom(x = 1, # number of successes
       size = 1, # size parameter
       prob = 0.5) # probability of success
```
```{r, echo = F, fig.height = 5, fig.width = 8, fig.align = "center"}
n <- 1
P <- 0.5
data <- dbinom(x=0:n,size=n, prob=P)
names(data) <- 0:n
cols <- rep("grey", n + 1)
cols[2] <- "tomato"
barplot(data, col = cols, main = "x = 1, n = 1, p = 0.5")
```
---

#### Calculating probabilities for a binomial distribution

The `dbinom` command gives the **probability mass function**; `dbinom` calculates the probability of observing `x` successes out of `size` independent trials, each with probability of success `prob`:

```{r}
dbinom(x = 5, # number of successes
       size = 10, # size parameter
       prob = 0.5) # probability of success
```
```{r, echo = F, fig.height = 5, fig.width = 8, fig.align = "center"}
n <- 10
P <- 0.5
data <- dbinom(x=0:n,size=n, prob=P)
names(data) <- 0:n
cols <- rep("grey", n + 1)
cols[6] <- "tomato"
barplot(data, col = cols, main = "n = 10, p = 0.5")
```
---
#### Calculating probabilities for a binomial distribution

The `dbinom` command gives the **probability mass function**; `dbinom` calculates the probability of observing `x` successes out of `size` independent trials, each with probability of success `prob`:

```{r}
dbinom(x = 3, # number of successes
       size = 10, # size parameter
       prob = 0.5) # probability of success
```
```{r, echo = F, fig.height = 5, fig.width = 8, fig.align = "center"}
n <- 10
P <- 0.5
data <- dbinom(x=0:n,size=n, prob=P)
names(data) <- 0:n
cols <- rep("grey", n + 1)
cols[4] <- "tomato"
barplot(data, col = cols, main = "n = 10, p = 0.5")
```
---
class: middle

A **cumulative distribution function** $F$ **also** describes the **distribution** of a random variable.

In particular, for a random variable $X$ with cumulative distribution function $F$,

$F(t)$ describes the probability of observing a value for $X$ that is less than or equal to $t$: 


$$F(t) = P(X\leq t)$$
---
#### Calculating probabilities for a binomial distribution

The `pbinom` command gives the **cumulative distribution function**; `pbinom` calculates the probability of observing `q` **or fewer** successes out of `size` independent trials, each with probability of success `prob`:

```{r}
pbinom(q = 1, # number of successes
       size = 1, # size parameter
       prob = 0.5) # probability of success
```
```{r, echo = F, fig.height = 5, fig.width = 8, fig.align = "center"}
n <- 1
P <- 0.5
data <- dbinom(x=0:n,size=n, prob=P)
names(data) <- 0:n
cols <- rep("grey", n + 1)
cols[1:2] <- "tomato"
barplot(data, col = cols, main = "q = 1, n = 1, p = 0.5")
```
---

#### Calculating probabilities for a binomial distribution

The `pbinom` command gives the **cumulative distribution function**; `pbinom` calculates the probability of observing `q` **or fewer** successes out of `size` independent trials, each with probability of success `prob`:

```{r}
pbinom(q = 5, # number of successes
       size = 10, # size parameter
       prob = 0.5) # probability of success
```
```{r, echo = F, fig.height = 5, fig.width = 8, fig.align = "center"}
n <- 10
P <- 0.5
data <- dbinom(x=0:n,size=n, prob=P)
names(data) <- 0:n
cols <- rep("grey", n + 1)
cols[1:6] <- "tomato"
barplot(data, col = cols, main = "q = 5, n = 10, p = 0.5")
```


---
class: middle

#### Sampling from a normal distribution

The `rnorm` command generates samples from a normal distribution:

```{r}
rnorm(10)
```

What are the mean and standard deviation of this normal distribution?

---
class: middle

#### Sampling from a normal distribution

The `rnorm` command generates samples from a normal distribution. By default the mean is 0 and the standard deviation is 1, but we can change these settings (also called arguments):

```{r}
rnorm(10, mean = 5, sd = 2)
```

What are the mean and standard deviation of this normal distribution?

---
class: middle

#### Sampling from a normal distribution

The `rnorm` command generates samples from a normal distribution. By default the mean is 0 and the standard deviation is 1, but we can change these settings (also called arguments):

```{r}
rnorm(n = 20, mean = 10, sd = 10)
```

---
#### Working with densities for a normal distribution

The `dnorm` command gives the **density function**;

```{r}
dnorm(1)
```
```{r, echo = F, fig.height = 5, fig.width = 8, fig.align = "center"}
x <- seq(-4, 4, length.out = 1000)
y <- dnorm(x)
plot(x, y, type = 'l', frame.plot = F,  ylab = '',
     main = "")
points(1, dnorm(1), col = "lightseagreen", cex = 2, pch =16)
```
---
#### Working with densities for a normal distribution

The `dnorm` command gives the **density function**;

```{r}
dnorm(1300, mean = 1100, sd = 200)
```
```{r, echo = F, fig.height = 5, fig.width = 8, fig.align = "center"}
x <- seq(300, 1900, length.out = 1000)
y <- dnorm(x, mean = 1100, sd = 200)
plot(x, y, type = 'l', frame.plot = F,  ylab = '',
     main = "")
points(1300, dnorm(1300,  mean = 1100, sd = 200), col = "lightseagreen", cex = 2, pch =16)
```
---

#### Calculating probabilities for a normal distribution

The `pnorm` command gives the **cumulative distribution function**; `pnorm` calculates the probability of observing a value of `q` **or less** for your normal random variable

```{r}
pnorm(q = 1)
```
```{r, echo = F, fig.height = 5, fig.width = 8, fig.align = "center"}
x <- seq(-4, 4, length.out = 1000)
y <- dnorm(x)
plot(x, y, type = 'l', frame.plot = F, axes = FALSE, ylab = '',
     main = "P(z < 1)")
axis(1, col = NA, col.ticks = 1)
polygon(c(-10, x[x<1], 1),  c(0, y[x<1], 0),
        col="lightseagreen", border = NA)
```
---
#### Calculating probabilities for a normal distribution

The `pnorm` command gives the **cumulative distribution function**; `pnorm` calculates the probability of observing a value of `q` **or less** for your normal random variable

```{r}
pnorm(q = 1300, mean = 1100, sd = 200)
```
```{r, echo = F, fig.height = 5, fig.width = 8, fig.align = "center"}
x <- seq(300, 1900, length.out = 1000)
y <- dnorm(x, mean = 1100, sd = 200)
plot(x, y, type = 'l', frame.plot = F, axes = FALSE, ylab = '',
     main = "P(x < 1300)")
axis(1, col = NA, col.ticks = 1)
polygon(c(300, x[x<1300], 1300),  c(0, y[x<1300], 0),
        col="lightseagreen", border = NA)
```
---
class: middle

Finally, for this week's lab, we'll be taking a look at data on rain in Seattle.

First, we need to load our dataset, which you can find [here](/data/seattleWeather_1948-2017.csv). We also load the `tidyverse` and `lubridate` package to help us with data cleaning. If you do not have the `lubridate` package, you may need to run `install.packages("lubridate")` first.

```{r load_data, warning = F, message = F, eval = F}
# load package of helpful datasets and functions
library(tidyverse)
library(lubridate)
# load data
daily_rain <-
  readr::read_csv("PATH_TO_DATA/seattleWeather_1948-2017.csv") 
```
```{r load_data_real, warning = F, message = F, echo = F}
# load package of helpful datasets and functions
library(tidyverse)
library(lubridate)
# load data
daily_rain <-
  readr::read_csv("../data/seattleWeather_1948-2017.csv") 
```
---
class: middle

Next, we clean our dataset by computing the `year` and `month` variables, and then calculating the days of rain in each month:

```{r}
daily_rain <- daily_rain %>%
  mutate(year = year(mdy(DATE)), # get year
         month = month(mdy(DATE))) %>% # get month
  mutate(year = ifelse(year > 2017, year - 100, year))
  # replace 2048, 2049, ... with 1948, 1949

monthly_rain <- daily_rain %>%
  group_by(month, year) %>% # group by month and year
  summarize(days_of_rain = sum(RAIN)) # count number of days of rain
```
---
class: middle

Finally, as a check, we visualize our data and make sure everything makes sense.

```{r, warning = F, fig.height = 6, fig.width = 6, fig.align = "center"}
ggplot(monthly_rain) + # plot monthly rain data
  geom_boxplot(aes(x = days_of_rain, # x axis variable
                   y = as.factor(month), # y axis variable
                   group = as.factor(month))) + # grouping variable
  xlab("Days of Rain") + ylab("Month")
```
