---
title: "Lab 03"
subtitle: "*Distributions*"
author: ""
institute: ""
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["style.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
 
---     
class: middle

In this lab, we'll practice working with different distributions in `R`.


---
class: middle

---
class: middle

First, we need to load our dataset, which you can find [here](/data/seattleWeather_1948-2017.csv). t We also load the `tidyverse` and `lubridate` package to help us with data cleaning. If you do not have the `lubridate` package, you may need to run `install.packages("lubridate")` first.

```{r load_data, warning = F, message = F}
# load package of helpful datasets and functions
library(tidyverse)
library(lubridate)
# load data
daily_rain <-
  readr::read_csv("../data/seattleWeather_1948-2017.csv") 
```
---
```{r}
daily_rain <- daily_rain %>%
  mutate(year = year(mdy(DATE)), # get year
         month = month(mdy(DATE))) %>% # get month
  mutate(year = ifelse(year > 2017, year - 100, year))
  # replace 2048, 2049, ... with 1948, 1949

monthly_rain <- daily_rain %>%
  group_by(month, year) %>% # group by month and year
  summarize(days_of_rain = sum(RAIN)) # count number of days of rain
```
---
```{r, warning = F, fig.height = 5, fig.width = 5, fig.align = "center"}
ggplot(monthly_rain) + 
  geom_boxplot(aes(x = days_of_rain,
                   y = as.factor(month), 
                   group = as.factor(month))) +
  xlab("Days of Rain") + ylab("Month")
```
---