---
title: "Lab 04"
subtitle: "*Sampling Distributions*"
author: ""
institute: ""
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["style.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
 
---     
class: middle

In this lab, we'll practice working with different distributions in `R`.\pause

First, let's take a look at binomial distributions.

---
class: middle

#### Sampling from a binomial distribution

Recall that every binomial distribution has a *size* parameter which is the number of trails and a *probability of success* parameter. Below, we sample ten coin flips using the `rbinom` function

```{r}
coin_flips <- rbinom(n = 10, # number of samples
                     size = 1, # size parameter
                     prob = 0.5) # probability of success
coin_flips
```

---
class: middle

#### Sampling from a binomial distribution

Now, consider an experiment where we flip a coin twenty times. The below code runs this experiment one thousand times for a fair coin and then for an unfair coin.

```{r, cache = T}
fair_flips <- rbinom(n = 1000, # number of experiments
                     size = 20, # size parameter
                     prob = 0.5) # probability of success
unfair_flips <- rbinom(n = 1000, # number of experiments
                       size = 20, # size parameter
                       prob = 0.6) # probability of success
```

---
class: middle

#### Sampling from a binomial distribution

Now, let's compare our results:

```{r, fig.height = 5, fig.width = 8, fig.align = "center"}
par(mfrow = c(1, 2)) # combine plots into one row with two columns
hist(fair_flips, main = "Fair", xlim = c(0, 20), breaks = 10)
hist(unfair_flips, main = "Unfair", xlim = c(0, 20), breaks = 10)
```

---
#### Calculating probabilities for a binomial distribution

The `dbinom` command gives the **probability mass function**; `dbinom` calculates the probability of observing `x` successes out of `size` independent trials, each with probability of success `prob`:

```{r}
dbinom(x = 1, # number of successes
       size = 1, # size parameter
       prob = 0.5) # probability of success
```
```{r, echo = F, fig.height = 5, fig.width = 8, fig.align = "center"}
n <- 1
P <- 0.5
data <- dbinom(x=0:n,size=n, prob=P)
names(data) <- 0:n
cols <- rep("grey", n + 1)
cols[2] <- "tomato"
barplot(data, col = cols, main = "x = 1, n = 1, p = 0.5")
```
---

#### Calculating probabilities for a binomial distribution

The `dbinom` command gives the **probability mass function**; `dbinom` calculates the probability of observing `x` successes out of `size` independent trials, each with probability of success `prob`:

```{r}
dbinom(x = 5, # number of successes
       size = 10, # size parameter
       prob = 0.5) # probability of success
```
```{r, echo = F, fig.height = 5, fig.width = 8, fig.align = "center"}
n <- 10
P <- 0.5
data <- dbinom(x=0:n,size=n, prob=P)
names(data) <- 0:n
cols <- rep("grey", n + 1)
cols[6] <- "tomato"
barplot(data, col = cols, main = "x = 1, n = 10, p = 0.5")
```
---

#### Calculating probabilities for a binomial distribution

The `pbinom` command gives the **distribution function**; `pbinom` calculates the probability of observing `q` **or fewer** successes out of `size` independent trials, each with probability of success `prob`:

```{r}
pbinom(q = 1, # number of successes
       size = 1, # size parameter
       prob = 0.5) # probability of success
```
```{r, echo = F, fig.height = 5, fig.width = 8, fig.align = "center"}
n <- 1
P <- 0.5
data <- dbinom(x=0:n,size=n, prob=P)
names(data) <- 0:n
cols <- rep("grey", n + 1)
cols[1:2] <- "tomato"
barplot(data, col = cols, main = "q = 1, n = 1, p = 0.5")
```
---

#### Calculating probabilities for a binomial distribution

The `pbinom` command gives the **distribution function**; `pbinom` calculates the probability of observing `q` **or fewer** successes out of `size` independent trials, each with probability of success `prob`:

```{r}
pbinom(q = 5, # number of successes
       size = 10, # size parameter
       prob = 0.5) # probability of success
```
```{r, echo = F, fig.height = 5, fig.width = 8, fig.align = "center"}
n <- 10
P <- 0.5
data <- dbinom(x=0:n,size=n, prob=P)
names(data) <- 0:n
cols <- rep("grey", n + 1)
cols[1:6] <- "tomato"
barplot(data, col = cols, main = "q = 5, n = 10, p = 0.5")
```


---
class: middle

First, we need to load our dataset, which you can find [here](/data/seattleWeather_1948-2017.csv). t We also load the `tidyverse` and `lubridate` package to help us with data cleaning. If you do not have the `lubridate` package, you may need to run `install.packages("lubridate")` first.

```{r load_data, warning = F, message = F}
# load package of helpful datasets and functions
library(tidyverse)
library(lubridate)
# load data
daily_rain <-
  readr::read_csv("../data/seattleWeather_1948-2017.csv") 
```
---
class: middle

Next, we clean our dataset by computing the `year` and `month` variables, and then calculating the days of rain in each month:

```{r}
daily_rain <- daily_rain %>%
  mutate(year = year(mdy(DATE)), # get year
         month = month(mdy(DATE))) %>% # get month
  mutate(year = ifelse(year > 2017, year - 100, year))
  # replace 2048, 2049, ... with 1948, 1949

monthly_rain <- daily_rain %>%
  group_by(month, year) %>% # group by month and year
  summarize(days_of_rain = sum(RAIN)) # count number of days of rain
```
---
class: middle

Finally, as a check, we visualize our data and make sure everything makes sense.

```{r, warning = F, fig.height = 6, fig.width = 6, fig.align = "center"}
ggplot(monthly_rain) + # plot monthly rain data
  geom_boxplot(aes(x = days_of_rain, # x axis variable
                   y = as.factor(month), # y axis variable
                   group = as.factor(month))) + # grouping variable
  xlab("Days of Rain") + ylab("Month")
```
---