---
title: "Lab 06"
subtitle: "*Correlation and Regression*"
author: ""
institute: ""
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["style.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
 
---     
class: middle

In this lab, we'll practice studying relationships between variables in `R`.

```{r, echo = F}
library(tidyverse)
```
---
class: middle

#### Flights data

Today, we'll look at data on flights leaving from New York for July 30, 2013.

Load the data using the following commands:

```{r, eval = F}
install.packages('nycflights13')
```

```{r}
flights <- nycflights13::flights %>% filter(month == 7 & day == 30)
```

---
class: middle

#### Visualizing the data

```{r, warning = F, fig.height = 4.5, fig.align = "center"}
ggplot(data = flights) +
  geom_point(aes(x = distance, y = air_time), col = "seagreen") + 
  xlab("Distance") + ylab("Air time") +
  ggtitle("Flights departing NYC (7/30/2013)") +
  theme_minimal()
```
---
class: middle

#### Visualizing the data

```{r, warning = F, fig.height = 4.5, fig.align = "center"}
ggplot(data = flights) +
  geom_point(aes(x = distance, y = arr_delay), col = "dodgerblue") + 
  xlab("Distance") + ylab("Minutes delayed") +
  ggtitle("Flights departing NYC (7/30/2013)") +
  theme_minimal()
```

---

class: middle

#### Computing correlation

```{r, fig.height = 4.5, fig.align = "center"}
cor(flights$distance, flights$air_time)
```
```{r, warning = F}
cor(flights$distance, flights$air_time,
    use = "complete.obs")
```
```{r, warning = F}
cor(flights$air_time, flights$distance,
    use = "complete.obs")
```
---
class: middle

#### Computing correlation

```{r, warning = F}
cor(flights$distance, flights$arr_delay,
    use = "complete.obs")
```
---
class: middle

#### Linear regression by "hand"

Suppose we want to use a linear function to predict air time from distance:

$$\widehat{y}_{air time}=b_0+b_1x_{distance}$$

**First, we compute the sample correlation $r$:**

```{r, warning = F} 
flights <- flights %>%
  filter(complete.cases(flights))
r <- cor(flights$distance, flights$air_time,
         use = "complete.obs")
r
```


---
class: middle

#### Linear regression by "hand"

Suppose we want to use a linear function to predict air time from distance:

$$\widehat{y}_{air time}=b_0+b_1x_{distance}$$


**Next, we estimate the slope:**

$$\widehat{b}_1=r\frac{s_y}{s_x}$$
```{r, warning = F}
s_y <- sd(flights$air_time, na.rm = T)
s_x <- sd(flights$distance, na.rm = T)
b_1_hat <- r * s_y / s_x
b_1_hat
```

---
class: middle

#### Linear regression by "hand"

Suppose we want to use a linear function to predict air time from distance:

$$\widehat{y}_{air time}=b_0+b_1x_{distance}$$


**Finally, we estimate the intercept, using the fact that the least squares line goes through the point $(\overline{x},\overline{y})$:**

$$\widehat{b}_0=\overline{y}-\widehat{b}_1\overline{x}$$
```{r, warning = F}
mean_y <- mean(flights$air_time, na.rm = T)
mean_x <- mean(flights$distance, na.rm = T)
b_0_hat <- mean_y - b_1_hat * mean_x
b_0_hat
```
---
class: middle

#### Linear regression by "hand"

Thus, our estimated least squares line is given by 
$$\widehat{y}_{air time}=\widehat{b}_0+\widehat{b}_1x_{distance}$$ with

$\widehat{b}_0=$ `r b_0_hat` and $\widehat{b}_1=$ `r b_1_hat`
---
class: middle

#### Linear regression by "hand"

```{r, warning = F, fig.height = 4.5, fig.align = "center", eval = F}
ggplot(data = flights) +
  geom_point(aes(x = distance, y = air_time), col = "seagreen") + 
{{ geom_abline(slope = b_1_hat, intercept = b_0_hat) + }}
  xlab("Distance") + ylab("Air time") +
  ggtitle("Flights departing NYC (7/30/2013)") +
  theme_minimal()
```
```{r, warning = F, fig.height = 4.5, fig.align = "center", echo = F}
ggplot(data = flights) +
  geom_point(aes(x = distance, y = air_time), col = "seagreen") + 
  geom_abline(slope = b_1_hat, intercept = b_0_hat) + 
  xlab("Distance") + ylab("Air time") +
  ggtitle("Flights departing NYC (7/30/2013)") +
  theme_minimal()
```

---
class: middle
#### Linear regression in R

```{r, warning = F, eval = F}
linear_reg_res <- lm(air_time~distance, data = flights)
summary(linear_reg_res)
```
```{r, warning = F, eval = F}
Call:
lm(formula = air_time ~ distance, data = flights)

Residuals:
    Min      1Q  Median      3Q     Max 
-45.459  -6.022   0.334   5.521  33.771 

Coefficients:
             Estimate Std. Error t value Pr(>|t|)    
*(Intercept) 1.218e+01  4.730e-01   25.76   <2e-16 ***
*distance    1.276e-01  3.657e-04  348.92   <2e-16 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 8.565 on 984 degrees of freedom
Multiple R-squared:  0.992,	Adjusted R-squared:  0.992 
F-statistic: 1.217e+05 on 1 and 984 DF,  p-value: < 2.2e-16
```
---
class: middle
#### Linear regression in R

```{r, warning = F}
linear_reg_res <- lm(air_time~distance, data = flights)
linear_reg_res$coefficients
c(b_0_hat, b_1_hat) # what we got by hand
```
---
class: middle
#### Linear regression in R: results

```{r, warning = F, eval = F}
linear_reg_res <- lm(air_time~distance, data = flights)
summary(linear_reg_res)
```
```{r, warning = F, eval = F}
Call:
lm(formula = air_time ~ distance, data = flights)

Residuals:
    Min      1Q  Median      3Q     Max 
-45.459  -6.022   0.334   5.521  33.771 

Coefficients:
             Estimate Std. Error t value Pr(>|t|)    
(Intercept) 1.218e+01  4.730e-01   25.76   <2e-16 ***
distance    1.276e-01  3.657e-04  348.92   <2e-16 ***
---
*Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 8.565 on 984 degrees of freedom
Multiple R-squared:  0.992,	Adjusted R-squared:  0.992 
F-statistic: 1.217e+05 on 1 and 984 DF,  p-value: < 2.2e-16
```
---
class: middle
#### Linear regression in R: results

```{r, warning = F, eval = F}
linear_reg_res <- lm(air_time~distance, data = flights)
summary(linear_reg_res)
```
```{r, warning = F, eval = F}
Call:
lm(formula = air_time ~ distance, data = flights)

Residuals:
    Min      1Q  Median      3Q     Max 
-45.459  -6.022   0.334   5.521  33.771 

Coefficients:
             Estimate Std. Error t value Pr(>|t|)    
(Intercept) 1.218e+01  4.730e-01   25.76   <2e-16 ***
distance    1.276e-01  3.657e-04  348.92   <2e-16 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 8.565 on 984 degrees of freedom
{{Multiple R-squared:  0.992,	Adjusted R-squared:  0.992}}
F-statistic: 1.217e+05 on 1 and 984 DF,  p-value: < 2.2e-16
```
---
class: middle

#### Interpreting results


**SLOPE:**

The slope is the estimated expected difference in $y$ for one observation is  if $x$ happened to be one unit larger. 

**INTERCEPT:**

The intercept describes the estimated expected value of $y$  if $x = 0$ \textbf{and} the linear model is valid  at $x = 0$ (often not true)

---
class: middle

#### Predictions

Suppose we want to predict the airtime for a flight with distance 3500 mi.

```{r, warning = F}
new_x <- 3500
predicted_y <- b_0_hat + b_1_hat * new_x
predicted_y
```

```{r, warning = F}
predict(linear_reg_res, newdata = data.frame(distance = new_x))
```
---
class: middle

#### Residual plot

Do you see any problems or outliers?

```{r, warning = F}
ggplot(data = flights) + 
  geom_point(aes(x = distance, y = resid(linear_reg_res))) +
  xlab("Distance") + ylab("Residual")+
  ggtitle("Residual plot") + theme_minimal()
```
---
class: middle

**In-class practice**: Repeat these steps except using distance to predict **arrival delays**.

---