---
title: "Lab 05"
subtitle: "*Sampling Distributions*"
author: ""
institute: ""
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["style.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
 
---     
class: middle

In this lab, we'll practice working with sampling distributions in `R`.\pause

---
class: middle

#### Sampling distributions

**Remember**: the **sampling distribution** is the theoretical distribution of a sample statistic.

Consider the following experiment. We flip a fair coin 100 times and record the proportion of flips where we observe heads, $\widehat{p}$. 

Here $\widehat{p}$ is a **sample statistic**. Each time we repeat the experiment, we obtain a new value, so we can model $\widehat{p}$ as a random variable.

The code below performs this experiment once.

```{r}
coin_flip <- rbinom(n = 1, # number of samples
                    size = 100, # size parameter
                    prob = 0.5) # probability of success
coin_flip
p_hat <- coin_flip / 100
p_hat
```

---
class: middle

#### Sampling distributions

By changing $n$ we can repeat the experiment multiple times.

```{r}
coin_flips <- rbinom(n = 10, # number of samples
                    size = 100, # size parameter
                    prob = 0.5) # probability of success
coin_flips
p_hats <- coin_flips / 100
p_hats
```

---

class: middle

Suppose we run this experiment once.

```{r}
set.seed(123)
coin_flip <- rbinom(n = 1, # number of samples
                     size = 100, # size parameter
                     prob = 0.5) # probability of success
coin_flip
p_hat <- coin_flip / 100
p_hat
```

---

class: middle

What is the **expected value** of $\widehat{p}$?

Compare with our observed value:

```{r}
p_hat
```
---
class: middle

What is the **standard error** of $\widehat{p}$?

$$SE_{\widehat{p}} = \sqrt{\frac{p(1-p)}{n}}$$

```{r}
true_se <- sqrt(0.5 * 0.5 / 100)
true_se
```

If we pretend we don't know $p$, we would **estimate** the standard error:

$$\widehat{SE}_{\widehat{p}} = \sqrt{\frac{\widehat{p}(1-\widehat{p})}{n}}$$

```{r}
estimated_se <- sqrt(p_hat * (1 - p_hat) / 100)
estimated_se
```
---
class: middle

Based on the theory from class, let's compute a 95% confidence interval based on $\widehat{p}$.

```{r}
my_first_ci <- c(p_hat - 2 * true_se, p_hat + 2 * true_se)
my_first_ci
```

Of course, we don't really know $p$ usually:

```{r}
more_realistic_ci <- c(p_hat - 2 * estimated_se,
                       p_hat + 2 * estimated_se)
more_realistic_ci 
```
---
class: middle

Let's also compute a 90% confidence interval based on $\widehat{p}$.

```{r}
z <- qnorm((1 + 0.9) / 2) # how do we derive this??
my_second_ci <- c(p_hat - z * true_se, p_hat + z * true_se)
my_second_ci
```

Compare with an 80% confidence interval based on $\widehat{p}$.

```{r}
z <- qnorm((1 + 0.8) / 2)
my_third_ci <- c(p_hat - z * true_se, p_hat + z * true_se)
my_third_ci
```
---
class: middle

#### Sampling distributions

Let's run this experiment 1000 times. This yields an **empirical sampling distribution**.

```{r, fig.width = 5, fig.height = 3.5, fig.align="center"}
set.seed(23)
coin_flips <- rbinom(n = 1000, # number of samples
                     size = 100, # size parameter
                     prob = 0.5) # probability of success
p_hats <- coin_flips / 100
hist(p_hats, main = "Empirical sampling distribution of p_hat")
```

---
class: middle

What is the **expected value** of $\widehat{p}$?

Compare with the mean of our empirical sampling distribution:

```{r}
mean(p_hats)
```

---
class: middle

What is the **standard error** of $\widehat{p}$?

$$SE_{\widehat{p}} = \sqrt{\frac{p(1-p)}{n}}$$

```{r}
sqrt(0.5 * 0.5 / 100)
```
Compare with the sample standard deviation of our empirical sampling distribution:

```{r}
sd(p_hats)
```
---
class: middle

For each $\widehat{p}$, we can compute a 95% confidence interval.

```{r, warning = F, message = F}
library(tidyverse)
set.seed(23)
ci_table <- data.frame(p_hat = p_hats) %>%
  mutate(true_se = sqrt(0.5 * 0.5 / 100),
         true_CI_lower = p_hat - 2 * true_se,
         true_CI_upper = p_hat + 2 * true_se)
head(ci_table)
```

---

class: middle

How many of our confidence intervals include the true $p$?

```{r, warning = F, message = F}
mean(ci_table$true_CI_lower < 0.5 & ci_table$true_CI_upper > 0.5)
```

---

class: middle

Of course, we don't know the true $p$! Let's repeat this analysis with estimated standard errors.


```{r, warning = F, message = F}
library(tidyverse)
ci_table <- data.frame(p_hat = p_hats) %>%
  mutate(est_se = sqrt(p_hat * (1 - p_hat) / 100),
         est_CI_lower = p_hat - 2 * est_se,
         est_CI_upper = p_hat + 2 * est_se)
head(ci_table)
mean(ci_table$est_CI_lower < 0.5 & ci_table$est_CI_upper > 0.5)
```

So it ends up not being a big deal when we have many flips.
---
class: middle

This week, we'll take a look at some data from the General Social Survey, which asks individuals about their life circumstances.

First, we need to load our dataset using the below code. We also load the `tidyverse` and `forcats` package to help us with data cleaning. If you do not have the `forcats` package, you may need to run `install.packages("forcats")` first.

```{r, message = F, warning = F}
library(lubridate)
library(forcats)

# load GSS data
data(gss_cat)

# filter to only the most recent data
gss_cat_2014 <- gss_cat %>%
  dplyr::filter(year == 2014)
```
---
class: middle

As always, we should do some quick exploratory plots and summaries.

```{r, message = F, warning = F}
summary(gss_cat_2014)
```
---
class: middle

As always, we should do some quick exploratory plots and summaries.

```{r, message = F, warning = F}
hist(gss_cat_2014$tvhours, main = "Hours of TV watched",
     xlab = "TV Hours")
```
---
class: middle

As always, we should do some quick exploratory plots and summaries.

```{r, message = F, warning = F}
table(gss_cat_2014$marital)
```
---
class: middle

In this week's lab, we'll also make use of some Tidy Tuesday data on all dogs that were adoptable on PetFinder.org on September 20, 2019. You can load it using this code:

```{r, message = F, warning = F, cache = T}
dog_descriptions <- 
  readr::read_csv(paste0('https://raw.githubusercontent.com/',
                         'rfordatascience/tidytuesday/master/',
                         'data/2019/2019-12-17/dog_descriptions.csv'))
```

---
class: middle

As always, we should do some quick exploratory plots and summaries.

```{r, message = F, warning = F}
summary(dog_descriptions)
```
---
class: middle

As always, we should do some quick exploratory plots and summaries.

```{r, message = F, warning = F}
table(dog_descriptions$breed_primary)
```
---